---
title: "qPCR Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(pander)

# Open the original data frame
file_data_in <- '2024-09-11_data_v2.csv'
dat1 <- read.csv(file_data_in)
dat1$sample_id <- paste(dat1$sample,dat1$condition,dat1$replicate,sep = '_')

# Make a data frame that averages across the qPCR replicates.
dat2 <- dat1
dat2$remove <- F
for(ii in 1:nrow(dat2)){
  tt <- dat2$sample_id[ii]
  t1 <- subset(dat2,dat2$sample_id == tt)
  if(dat2$replicate_qpcr[ii] == 'r1'){dat2$ct[ii] <- mean(t1$ct)}else{dat2$remove[ii] <- T}
}
dat2 <- subset(dat2,dat2$remove == F)

# Determine the standard curve for the standards.
std1 <- subset(dat2,dat2$sample == 'standard')
std1$copies <- std1$standard_quantity/std1$dilution_factor
std1$lncopies <- log10(std1$copies)

plot(std1$lncopies, std1$ct)


# Perform linear regression
model <- lm(ct ~ lncopies, data = std1)

# Get the formula and R^2 value from the model
formula <- gsub('x','log10(ct)', paste("y = ", round(coef(model)[2], 2), "*x +", round(coef(model)[1], 2)))
r_squared <- summary(model)$r.squared
r_squared_text <- paste("R² = ", round(r_squared, 3))

# Create the plot
p_std <- ggplot(std1, aes(x = lncopies, y = ct)) +
  geom_point() +                         # Scatter plot
  geom_smooth(method = "lm", col = "red") +  # Linear regression line
  labs(title = paste(formula, "\n", r_squared_text),   # Title with formula and R²
       x = "Standard Dilution",          # X-axis label
       y = "Ct") +                       # Y-axis label
  theme_classic()                        # A minimal theme

# Process the different samples
dat3 <- subset(dat2,dat2$sample != 'standard')
# Get model coefficients
intercept <- coef(model)[1]
slope <- coef(model)[2]

# Calculate lncopies for dat3 based on the model
dat3$lncopies <- (dat3$ct - intercept) / slope
dat3$copies <- 10^dat3$lncopies

# Plot that shows all the data from the samples
plot_ct <- ggplot(dat3, aes(x = sample, y = ct, fill = condition)) +
  geom_boxplot() +
  labs(title = "CT values by Sample and Condition",
       x = "Sample",
       y = "CT Value",
       fill = "Condition") +
  ylim(1, 40) +  # Set y-axis limits from 0 to 40
  theme_classic()


plot_copies <- ggplot(dat3, aes(x = sample, y = copies, fill = condition)) +
  geom_boxplot() +
  scale_y_log10() +  # Apply log10 scale to y-axis
  labs(title = "Copies by Sample and Condition",
       x = "Sample",
       y = "Copies (log10)",
       fill = "Condition") +
  theme_classic()

# Create vectors to store sample IDs and p-values
sample_ids <- c()
p_values <- c()

# Loop through each unique sample in dat3
for (sample_n in unique(dat3$sample)) {
  # Subset the data for the current sample
  sample_data <- subset(dat3, sample == sample_n)
  
  # Perform ANOVA
  model <- aov(copies ~ condition, data = sample_data)
  
  # Extract p-value from ANOVA summary
  p_value <- summary(model)[[1]]$`Pr(>F)`[1]
  
  # Append the sample ID and p-value to vectors
  sample_ids <- c(sample_ids, sample_n)
  p_values <- c(p_values, p_value)
}

# Create a data frame with sample IDs and p-values
anova_pvalues_df <- data.frame(sample = sample_ids, p_value = p_values)

# Display the data frame
anova_pvalues_df

# Save the processed data for viewing.
file_data_out <- gsub('.csv','_analysis.csv',file_data_in)
write.csv(dat3,file_data_out)

```
## Results

### Standard Curve
The standard curve was generated based on the known standard dilutions. The following plot represents the standard curve fitted with linear regression:

```{r plot_std, echo=FALSE, message=FALSE}
p_std
```

The equation for the standard curve is:

`r formula`


The R-squared value for the model is `r r_squared_text`.

\newpage 

### Sample Data Analysis

The following plots show the distribution of Ct values and the copy number of the samples under different conditions:

```{r plot_ct, echo=FALSE, fig.width=5, fig.height=3.5}
plot_ct
```

```{r plot_copies, echo=FALSE, fig.width=5, fig.height=3.5}
plot_copies
```

### ANOVA Results

An ANOVA was performed to test for significant differences between conditions for each sample. The following table summarizes the p-values:

```{r echo=FALSE, fig.cap="", warning = FALSE, message = FALSE}
pander(anova_pvalues_df, style = 'rmarkdown', justify = c('left', 'left'), caption = "ANOVA Results for Conditions by Sample")
```
